Nodal Discontinuous Galerkin Method for the Euler Equations in GR

### 2018-11-19 Documents/Euler/dgEulerGR.tex
Added to dgEulerGR.tex

\subsection{Example: $f\left(x\right)=x^{2}$}
We can show this directly for simple functions. We take as an example $f\left(x\right)=x^{2}$:
\begin{align}
    f\left(x\right)&=x^{2}=\left[x_{L}+\eta\left(x_{R}-x_{L}\right)\right]^{2}=x_{L}^{2}+\eta^{2}\left(x_{R}-x_{L}\right)^{2}+2\,x_{L}\,\eta\left(x_{R}-x_{L}\right)\\
    &=x_{L}^{2}+\eta^{2}\,x_{R}^{2}+\eta^{2}\,x_{L}^{2}-2\,\eta^{2}\,x_{L}\,x_{R}+2\,\eta\,x_{L}\,x_{R}-2\,\eta\,x_{L}^{2}.
\end{align}
Now we add and subtract the interpolated solution: $\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}$. This yields:
\begin{align}
    x^{2}&=x_{L}^{2}+\eta^{2}\,x_{R}^{2}+\eta^{2}\,x_{L}^{2}-2\,\eta^{2}\,x_{L}\,x_{R}+2\,\eta\,x_{L}\,x_{R}-2\,\eta\,x_{L}^{2}+\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}-\eta\,x_{H}^{2}-\left(1-\eta\right)x_{L}^{2}\\
    &=\left[\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}\right]+\left[-\eta\left(1-\eta\right)x_{L}^{2}-\eta\left(1-\eta\right)x_{H}^{2}+2\,\eta\,x_{L}\,x_{H}\left(1-\eta\right)\right]\\
    &=\left[\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}\right]-\eta\left(1-\eta\right)\left[x_{L}^{2}+x_{H}^{2}-2\,x_{L}\,x_{H}\right]\\
    &=\left[\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}\right]-\eta\left(1-\eta\right)\left(x_{H}-x_{L}\right)^{2}\\
    &=\left[\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}\right]-\eta\left(1-\eta\right)\left(\Delta x\right)^{2}.
\end{align}
The term in square brackets is the interpolated solution, and the second term is the remainder, i.e. the error in the interpolated solution. We show the results of this in the following figure:
\begin{figure}[H]
\centering
\includegraphics[scale=0.75]{LinearInterpolationError}
\end{figure}

\subsection{Mass Constant}
The mass constant, $C_{D}$, from the 3+1 GR hydrodynamics equations is:
\begin{equation}
    C_{D}\left(r\right)=\psi\left(r\right)^{6}\,\alpha\left(r\right)\,r^{2}\,\rho\left(r\right)\,W\left[v\left(r\right)\right]\,v\left(r\right).
\end{equation}
From the data we have this exact value at the two endpoints:
\begin{align}
    C_{D}\left(r_{L}\right)&\equiv C^{L}_{D}=\psi_{L}^{6}\,\alpha_{L}\,r_{L}^{2}\,\rho_{L}\,W_{L}\,v_{L}\\
    C_{D}\left(r_{H}\right)&\equiv C^{H}_{D}=\psi_{H}^{6}\,\alpha_{H}\,r_{H}^{2}\,\rho_{H}\,W_{H}\,v_{H}.
\end{align}

>>>>

\subsection{Example: $f\left(x\right)=x^{2}$}
We can show this directly for simple functions. We take as an example $f\left(x\right)=x^{2}$:
\begin{align}
    f\left(x\right)&=x^{2}=\left[x_{L}+\eta\left(x_{R}-x_{L}\right)\right]^{2}=x_{L}^{2}+\eta^{2}\left(x_{R}-x_{L}\right)^{2}+2\,x_{L}\,\eta\left(x_{R}-x_{L}\right)\\
    &=x_{L}^{2}+\eta^{2}\,x_{R}^{2}+\eta^{2}\,x_{L}^{2}-2\,\eta^{2}\,x_{L}\,x_{R}+2\,\eta\,x_{L}\,x_{R}-2\,\eta\,x_{L}^{2}.
\end{align}
Now we add and subtract the interpolated solution: $\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}$. This yields:
\begin{align}
    x^{2}&=x_{L}^{2}+\eta^{2}\,x_{R}^{2}+\eta^{2}\,x_{L}^{2}-2\,\eta^{2}\,x_{L}\,x_{R}+2\,\eta\,x_{L}\,x_{R}-2\,\eta\,x_{L}^{2}+\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}-\eta\,x_{H}^{2}-\left(1-\eta\right)x_{L}^{2}\\
    &=\left[\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}\right]+\left[-\eta\left(1-\eta\right)x_{L}^{2}-\eta\left(1-\eta\right)x_{H}^{2}+2\,\eta\,x_{L}\,x_{H}\left(1-\eta\right)\right]\\
    &=\left[\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}\right]-\eta\left(1-\eta\right)\left[x_{L}^{2}+x_{H}^{2}-2\,x_{L}\,x_{H}\right]\\
    &=\left[\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}\right]-\eta\left(1-\eta\right)\left(x_{H}-x_{L}\right)^{2}\\
    &=\left[\eta\,x_{H}^{2}+\left(1-\eta\right)x_{L}^{2}\right]-\eta\left(1-\eta\right)\left(\Delta x\right)^{2}.
\end{align}
The term in square brackets is the interpolated solution, and the second term is the remainder, i.e. the error in the interpolated solution. So we see that when using linear interpolation to interpolate the function $f\left(x\right)=x^{2}$ we introduce an error (of the order $\left(\Delta x\right)^{2}$) that goes to zero as the step-size goes to zero.

As a final step we replace $\eta$ with $x$, yielding:
\begin{align}
    f\left(x\right)=x^{2}&=x_{L}^{2}+\f{x_{H}^{2}-x_{L}^{2}}{x_{H}-x_{L}}\left(x-x_{L}\right)-\left(x-x_{L}\right)\left(x_{H}-x\right)\\
    &=f\left(x_{L}\right)+\f{\Delta f}{\Delta x}\left(x-x_{L}\right)-\left(x-x_{L}\right)\left(x_{H}-x\right).
\end{align}

We show the results of this in the following figure:
\begin{figure}[H]
\centering
\includegraphics[scale=0.75]{LinearInterpolationError}
\end{figure}

\subsection{Mass Constant}
The mass constant, $C_{D}$, from the 3+1 GR hydrodynamics equations is:
\begin{equation}
    C_{D}\left(r\right)=\psi\left(r\right)^{6}\,\alpha\left(r\right)\,r^{2}\,\rho\left(r\right)\,W\left[v\left(r\right)\right]\,v\left(r\right).
\end{equation}
When we map from the radial coordinates from the original data to the mesh in \texttt{thornado} we don't introduce any error, because it is a linear map. This means that all the error we introduce must be due to $\rho$, $W$, and $v$. So, when deriving the error we consider the different function:
\begin{equation}
    C'_{D}\left(r\right)\equiv\f{C_{D}\left(r\right)}{\psi\left(r\right)^{6}\,\alpha\left(r\right)\,r^{2}}=\rho\left(r\right)\,W\left[v\left(r\right)\right]\,v\left(r\right).
\end{equation}

From the data we have this exact value at the two endpoints:
\begin{align}
    C'_{D}\left(r_{L}\right)&\equiv C'^{L}_{D}=\rho_{L}\,W_{L}\,v_{L}\\
    C'_{D}\left(r_{H}\right)&\equiv C'^{H}_{D}=\rho_{H}\,W_{H}\,v_{H}.
\end{align}
